---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Esteshly Torres"
output: 
  html_document:
    theme: cerulean
    highlight: breezedark
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)
```

<br><br>

## Introduction
+ Tidy Tuesday Dataset: [Plastic Polution Dataset](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-26/readme.md)

I chose the Tidy Tuesday dataset that debuted on 01/26/2021 because it has to do with an audit of global plastic production and the cleanup efforts. This is in an attempt to understand where the most plastic production is occurring, where it is ending up and how we as a population can make positive changes for the environment.
```{r read-examine-data}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')
glimpse(plastics)
```

<br><br>

## Question 1

#### Which country had the most plastic polution clean up from 2019-2020?

### Methods
```{r Plot_and_Data_for_Question_1}
highest_grand_total_data<-plastics%>%
  select(grand_total,country)%>%
  group_by(country)%>%
  arrange(-grand_total)%>%
  summarize(total_poln=sum(grand_total))%>%
  arrange(-total_poln)

highest_grand_total_data

top_10_highest_poln_data<-
  highest_grand_total_data%>%
  filter(total_poln>21104)%>%
  filter(country!="EMPTY")%>%
  # consider using case_when
  mutate(country=ifelse(country=="Taiwan_ Republic of China (ROC)","Taiwan",country))%>%
  mutate(country=ifelse(country=="ECUADOR","Ecuador",country))%>%
  mutate(country=ifelse(country=="NIGERIA","Nigeria",country)) %>%
  group_by(country) %>%
  summarize(total_poln = sum(total_poln))
  
top_10_highest_poln_data

plot_quest_1<-top_10_highest_poln_data%>%
ggplot()+
  aes(x=fct_reorder(country,total_poln,.desc = TRUE),
      y=total_poln)+
  geom_col(fill="turquoise")+
  labs(x="Country",
       y="Total Polution Cleanup")+
  theme(axis.text.x=element_text(angle=25))+
  theme_minimal()

print(plot_quest_1)
```

### Answer

Taiwan had the highest amount of plastic pollution cleanup.


<br><br>

## Question 2

#### Which country had the most cleanup events from 2019-2020?

### Methods
```{r Plot_and_Data_for_Question_2}
num_event_by_country<-plastics%>%
  select(country,num_events)%>%#Selecting country and number of events columns
  group_by(country)%>%
  summarize(total_events=sum(num_events))%>% #Summing up events per country
  arrange(-total_events)%>%#Arrange data from greatest to least
  filter(country!="EMPTY")%>%#Getting rid of Empty aka N/A data
   mutate(country=ifelse(country=="NIGERIA","Nigeria",country)) %>%
 slice(1:11)%>%#Getting top 10 countries (2 for Nigeria)
  mutate(country=ifelse(country=="United States of America","USA",country))%>%
   group_by(country)%>% #Need to group again to group two Nigerian data events
  summarise(total_events=sum(total_events))%>% #Adding up Nigerian data events
  arrange(-total_events) #Arrange data from greatest to least

num_event_by_country

num_event_by_country%>%
  ggplot()+
  aes(x=fct_reorder(country,total_events,.desc = TRUE),#Ordered bars from greatest to least
      y=total_events)+
  geom_col(fill="lightseagreen")+
  labs(x="Country",y="Total Number of Counting Events",title = "Top Ten Countries: Greatest Number of Counting Events (2019-2020)")+
  theme_classic()

```

### Answer

The USA had the most cleanup events.

<br><br>

## Question 3

#### Which country had the most volunteers between 2019 and 2020?

### Methods
```{r Plot_and_Data_for_Question_3}
total_volunteers_data<-plastics%>%
  select(country,volunteers)%>%#Selecting volunteers and countries columns
  group_by(country)%>%
  summarize(total_volunteers=sum(volunteers))%>%#Adding up volunteer totals
  arrange(-total_volunteers)%>%#Arranging from greatest value to least
  filter(country!="EMPTY")%>%#Getting rid of Empty aka N/A data
  slice(1:11)%>%#Getting top 10 countries (2 for Nigeria)
  mutate(country=ifelse(country=="Taiwan_ Republic of China (ROC)",
                        "Taiwan",country))%>%#Country name change
  mutate(country=ifelse(country=="NIGERIA","Nigeria",country))%>% #Uniform name for Nigerian data
  mutate(country=ifelse(country=="United States of America","USA",country))%>%
   group_by(country)%>% #Need to group again to group two Nigerian data volunteers
  summarise(total_volunteers=sum(total_volunteers))%>% #Adding up Nigerian data volunteers
  arrange(-total_volunteers) #Arrange data from greatest to least

total_volunteers_data

total_volunteers_data%>%
  ggplot()+
  aes(x=fct_reorder(country,total_volunteers,.desc = TRUE),#Ordering bars from greatest to least
      y=total_volunteers)+
  geom_col(fill="deepskyblue4")+
  labs(x="Country",
       y="Total Number of Volunteers",
       title = "Top Ten Countries: Most Volunteers (2019-2020)")+
  theme_bw()
```

### Answer

Indonesia had the most volunteers.


<br><br>

## Question 4

#### Which country had the most high density polyethylene count (hdpe) from 2019-2020?

### Methods

```{r Plot_and_Data_for_Question_4}
plastic_hdpe_data<-plastics%>%
  select(country,hdpe)%>% #Selecting two specific columns
  group_by(country)%>%
  summarize(total_hdpe=sum(hdpe))%>% # Add up all of the hdpe totals for each country, creating new column
  arrange(-total_hdpe)%>% #Arrange data from greatest to least
  filter(country!="EMPTY")%>% #Getting rid of Empty aka N/A data
   slice(1:11)%>% #Getting top 10 countries (2 for Nigeria)
  mutate(country=ifelse(country=="NIGERIA","Nigeria",country))%>% # Uniform name for Nigerian data
  mutate(country=ifelse(country=="United States of America","USA",country))%>% #Changing name to USA
  group_by(country)%>% #Need to group again to group two Nigerian data hdpe
  summarise(total_hdpe=sum(total_hdpe))%>% #Adding up Nigerian data hdpe
  arrange(-total_hdpe) #Arrange data from greatest to least

plastic_hdpe_data

plastic_hdpe_data%>%
  ggplot()+
  aes(x=fct_reorder(country,total_hdpe,.desc = TRUE), #Ordering data in graph from greatest to least
      y=total_hdpe)+
  geom_col(fill="royalblue4")+
  labs(x="Country",y="Total HDPE Count",
       title = "Top Ten Countries: Greatest HDPE Plastic Cleanup (2019-2020)")+
  theme_light()

```

### Answer

Nigeria had the most hdpe plastic cleaned up.
